Cross-scale integration of knowledge for predicting species ranges: a metamodeling framework

### 2014-10-13 tex/manuscript.tex
First round of edits to the manuscript

%%% available colors:
%%% red, green, blue, cyan, magenta, yellow, gray, drakfray, lightgray, brown, lime, olive, orange, pink, purple, teal, violet
%
%\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
%\newcommand{\ib}[1]{\todo[color=magenta!40]{\textsuperscript{IB}#1}} 				
%\renewcommand{\aa}[1]{\todo[color=violet!40]{\textsuperscript{Aitor A}#1}}
\newcommand{\ia}[1]{\todo[color=blue!20]{\textsuperscript{Isabelle A}#1}}
\newcommand{\db}[1]{\todo[color=violet!20]{\textsuperscript{Dominique B}#1}}
%\newcommand{\ab}[1]{\todo[color=teal!40]{\textsuperscript{Alyssa B}#1}}
%\newcommand{\fd}[1]{\todo[color=orange!30]{\textsuperscript{Frédérik D}#1}}
%\newcommand{\rd}[1]{\todo[color=orange!40]{\textsuperscript{Ronnie D}#1}}
\newcommand{\mjf}[1]{\todo[color=green!60]{\textsuperscript{Marie-Josée F.}#1}}
%\newcommand{\tf}[1]{\todo[color=yellow!60]{\textsuperscript{Tony F.}#1}}
\newcommand{\dm}[1]{\todo[color=lime!50]{\textsuperscript{Dan M.}#1}}
\newcommand{\ks}[1]{\todo[color=green!40]{\textsuperscript{Kevin S}#1}}
%\newcommand{\ns}[1]{\todo[backgroundcolor=red!100, linecolor=orange!100, bordercolor=cyan!100]{{\color{magenta!40}\textsuperscript{Strigul}#1}}}
\newcommand{\wt}[1]{\todo[color=orange!80]{\textsuperscript{Wilfried T}#1}}
\newcommand{\dg}[1]{\todo[color=pink!40]{\textsuperscript{DG}#1}}

@@

%TC:break abstract
\linenumbers
\begin{abstract}
\ia{Some section of the manuscript are a bit long and may be shorterned. I have made some suggestion but I think it should be shorten more.}
\ia{The abstract would need a bit of work in the formulation and editing. I have made some suggestions.}
% limited to 200 words
There is great interest in modeling species ranges, resulting in a great diversity of modeling approaches.
\dm{There is great interest in modeling species ranges – both potential and actual, resulting in a great diversity of modeling approaches.}
However, in general, individual approaches include only a small subset of the total available information about a species. 
\ia{delete first two sentences, replace with: \\
Current worldwide preoccupations with species range shifts have resulted in a great diversity of approaches in species range modelling. Each approach typically includes only a small subset of the total available information about a species; correlative models customarily predict...}
Correlative models often predict presence or absence as a function of climate, but ignore  smaller-scale processes such as growth, fecundity, and dispersal.
\ia{see suggestion for a rewrite of the remainder:\\
What’s more, these approaches frequently result in divergent predictions, with no straightforward method to reconcile them. We present a flexible framework allowing the integration of models across multiple scales. Using hierarchical Bayesian methods, we produce ametamodel which, by incorporating all of the information used as input for the original scale-specific sub-models, produce probabilistic estimates of species presence andreports uncertainty from both data and process error. We provide two contrasting examples to illustrate our approach, and demonstrate that it can substantially reduce uncertainty when projecting beyond the range of the original data. Our method could be an invaluable tool for conservation biologists and land managers, despite the technical challenge of its actual implementation. We strongly encourage collaboration between modelers and practitioners in overcoming these hurdles and achieving greater integration of knowledge, a goal which we are confident will be of wide interest.
}
Furthermore, the diversity of approaches often results in divergent predictions for the same species, with no simple way to reconcile these predictions.
We present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods. 
The resulting metamodel produces probabilistic estimates of species presence, reports uncertainty from both data and process error from all sub-models, and incorporates all of the information used as input for the original scale-specific sub-models.
\ks{Change above to:\\
The resulting metamodel produces probabilistic estimates of species presence, reports uncertainty from both data and process error from all sub-models used, while also  synthesizing all of the information as an input for the original scale-specific sub-models.}
We illustrate our approach using two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of the original data.
We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
\dg{That’s the smartest comment of the whole paper!!!}
Although implementation of our method can be technically challenging, we anticipate the results will be of wide interest, and we encourage collaboration between modelers and practitioners in applying our framework.
\end{abstract}

@@

Models of species range limits have wide applications and can play a large role in conservation biology, where they can be used as decision-making tools in biodiversity management \citep{Guisan2013}.
\ia{This sentence is repetitive and somewhat contradictory with the last one of the paragraph. Suggest deletion}
Due to large temporal and spatial scales as well as the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the ecological processes generating species range limits \citep{Wu1995, Levin1998}. 
Hence, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, evolution \citep{Jay2012}, hybrid zone dynamics \citep{Engler2013}, and forecasting the impacts of climate change on species distributions \citep{Thuiller2011, Blois2013, Thuiller2014}. 
However, despite having recognized potential to support decision making, the contributions of these models to conservation and applied ecology remain limited \citep{Guisan2013}.

\ia{Miss one sentence here. To make a link with the sentence above, you should briefly state why models contribution remains limited. }
Two important constraints on modeling ecological processes can prevent models from making unbiased predictions with acceptable uncertainty: (1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
Another constraint is a paucity of data over wide ranges of ecological conditions and scales.
\dg{change to:\\
The paucity of data over theses ecological conditions and scales is a severe constraint. }
Poor data or incoherent models will result in predictions with unacceptably high uncertainty or precise but biased predictions.
In recent decades, however, modeling techniques have proliferated to take advantage of what is available in terms of both data and modeling platforms.
Model diversification also arises due to the diverse processes generating species ranges \citep{Soberon2007}.
\dg{Just to encourage self citation, this paper is more recent and supported by data, not only discussion: \\

 Boulangeat, I., Gravel, D. and Thuiller, W. 2012. Disentangling the underlying mechanisms of species abundance distribution using a comprehensive and nested modeling framework. Ecology Letters 15 : 584-593. }
In the context of forecasting, which is the focus here, mechanistic and correlative approaches represent quite different modes of inference, but both have strong limitations. 
\ia{delete "which is the focus here"}
Fine-scale mechanistic models may capture important ecological processes quite well, but due to their specificity, they may perform poorly when applied at the scale of species ranges.
For instance, biotic interactions (including trophic interactions) are usually not modeled mechanistically at the scale of species distributions because they are poorly known, or have not been recorded, despite being considered a key determinant of range limits \citep{Holt2005, Soberon2007, Pigot2013}. 
On the other hand, coarse-scale correlative models statistically relating species occurrences to associated distributions of other variables have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
However, their predictions rely on the constancy of the relationships between occurrences and explanatory variables in time and space, implying that the selected variables are related to the processes limiting species ranges and that their correlations are constant for calibration and projection ranges in space and time \citep{Dormann2007}. 
Extrapolating beyond the scope of the original data (e.g., predicting ranges based on future climate) is therefore problematic, because nonlinearities in responses to novel combinations of the explanatory variables cannot be accommodated in models that do not simulate the underlying processes.
In cases where a selected explanatory variable indirectly accounts for a process (e.g., where the strength of biotic interactions are correlated with an environmental variable), extrapolation will produce biased predictions in temporal and spatial domains where the process and the environmental variable are decoupled.
Furthermore, correlative models often do not explicitly incorporate ecological processes into their formulations, and when applied at coarse scales, they may miss important fine-scale processes, making a multi-scale approach desirable when accurate forecasting is required \citep{Thuiller2013}.
\ia{replace: \\
 desirable when accurate forecasting is required \\
with:
 more suitable}
On the other hand process-based models often require a variety of assumptions which may be poorly understood across space and time.
\ia{Finish abruptely here. Feels like one sentence is missing}
\ks{Feels like it is getting snuck in here…almost an after thought… I would recommend either removing it or developing it a tad further. }
%% ----------------------------------
%% ----------------------------------
%% ----------------------------------

\subsection*{Toward an integrated approach for modeling ranges}
To include knowledge of species from multiple sources and scales into distribution models, we present an application of hierarchical Bayesian methods that uses outputs from multiple model outputs to inform the results of the final model.
Our approach is an alternative to other approaches to multimodel inference that offers improved flexibility in incorporating data and theory from multiple scales and provides transparent estimation of uncertainty.
The challenge is to account for all this available knowledge while producing a single prediction, potentially mitigating the limitations inherent in each modeling approach and contributing to more robust predictions \citep{Pearson2003, Guisan2005, Araujo2006, Quillet2010}.
\ia{replace above sentence with: \\
``By producing a single prediction from multiple models, we aim at mitigating the limitations inherent in each modeling approach and contributing to more robust predictions ''}
\ks{change previous 2 sentences to:\\
Our approach succeeds in providing a much more throrough implementation of data from multiple models, while incorportating multiple scales and transparency in our estimation of uncertainty. A number of challenge(s) arised with this approach: (1) accounting for all the available data, (2) producing a single congruent prediction from all the models implemented, and (3) mitigating the potential limitations within each modeling approach }
Some researchers have explicitly reproduced the hierarchical nature of ecological systems using hierarchical models \citep[e.g.][]{Royale2008, Boulangeat2012, Catterall2012, Strigul2012, Soranno2014}. 
Recent literature has also focused on the development of hybrid models, which allow for flexible combinations between mechanistic and phenomenological sub-models \citep{Gallien2010, Thuiller2013}. 
For instance, correlative models are used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while mechanistic approaches can include space and time dynamics and therefore account for dispersal processes \citep{Kearney2008}. 
This complementarity has been used to merge the two model types into hybrid integrated models \citep[e.g.][]{Keith2008, Anderson2009, Smolik2010, Boulangeat2014}. 
\wt{Dormann et al. suggested to separate hybrid and integrated. 
Dormann, C.F., Schymanski, S.J., Cabral, J.S., Chuine, I., Graham, C.H., Hartig, F., Kearney, M., Morin, X., Römermann, C., Schröder, B. \& Singer, A. (2012) Correlation and process in species distribution models: bridging a dichotomy. Journal of Biogeography,
}
However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are not easy to test \citep{Gallien2010},
\wt{and hardly known}
and uncertainties are very difficult to quantify and identify from its many different sources. 
Moreover, hybrid approaches are not suitable for merging models that are based on the same processes but with different underlying hypotheses.

A second approach is to directly combine predictions.
\ia{A second approach recently explored by modelers is to directly combine predictions.}
If models operate at the same scale and have compatible parameterizations, predictions can be combined using multi-model inference \citep[e.g., model averaging, ensemble forecasting;][]{Thuiller2004, Araujo2007}. 
If models are not comparable because they are based on different data or hypotheses, the use of convergent predictions from multiple models is a way to provide more robust projections \citep{Morin2009, Marmion2009, Serra-Diaz2013}.
\wt{Cheaib, A., Badeau, V., Boe, J., Chuine, I., Delire, C., Dufrêne, E., François, C., Grittid, E.S., Legay, M., Pagé, C., Thuiller, W., Viovy, N. \& Leadley, P. (2012) Climate change impacts on tree ranges: model inter-comparison facilitates understanding and quantification of uncertainty. Ecology Letters, 15, 533-544.}
However, the applicability of ensemble forecasts is limited; for example, it is not currently possible to evaluate the effects of convergent predictions on the total uncertainty of the outcomes.
\ia{paragraph break here}
One of the greatest challenges inherent in both approaches is estimating uncertainty, which is of utmost importance in a prediction context.
A more comprehensive understanding of uncertainty can guide biodiversity management and prioritize future data collection by identifying parameters that make large contributions to the total uncertainty in the model predictions \citep{McMahon2011}. 
In models that encompass multiple scales or levels of organization, it is particularly challenging to evaluate all sources of uncertainty, especially those related to model specification (\citealt{Calder2003} but see \citealt{Conlisk2013}).
\dg{‘uncertainty’ is used 6 times over 5 sentences. Would help some rewording to limit redundancy}
Clearly, an approach that integrates predictions and uncertainty from multiple model types is needed \citep{Beck2012, Thuiller2013}. 
Such an approach would incorporate multiple modes of inference and integrate data from multiple sources and scales \citep{Levin1992, Peters2004, Thuiller2013}.

Here, we present a framework for modeling species range dynamics that aims to account for all the information available on a focal species, from a variety of sources and scales.
Contrary to hybrid and hierarchical models, the aim is not to link different sub-models into a single model, but to condition the predictions of a metamodel at the target scale (e.g., an entire species' range) with information from independent sub-models at a variety of spatial scales, allowing for as much flexibility as possible regarding the type of information included. 
We use the power and generality of a hierarchical Bayesian framework, which allows us to include multiple data sources and modes of inference \citep{VanOijen2005, Clark2006, Hobbs2011, Hartig2012}. 
Another advantage of Bayesian methods is that uncertainty in model outputs is intuitive to interpret and reflects uncertainty at all levels of organization \citep{Cressie2009, Hobbs2011}. 
We illustrate our approach with two examples.
First, we use simulated data to present a complete example of the application of the framework to multiple information sources that are relevant at very different scales.
In a second example, we apply the framework to combine predictions of a correlative model relating the range of sugar maple (\emph{Acer saccharum}), a widespread foundation tree species from eastern North America, to climate with the predictions of a phenological model, Phenofit, with the goal of improving uncertainty estimation when projecting model predictions to future climate.
\dg{ref missing for phenofit}

@@

The key idea of our approach is to constrain the parameters of a correlative metamodel using predictions made by one or several mechanistic sub-models from other spatial scales.
\wt{does not really have to, right? (wrt different scales)}
The role of the metamodel is to integrate data at the same ecological scale as predictions. 
As an example, the metamodel could take the form of a simple correlative \ac{SDM}.
Ideally, each sub-model incorporates different data types and hypotheses that specify underlying ecological mechanisms contributing to the model output.
The output of sub-models should provide additional predictions at a comparable ecological scale at which the metamodel operates, although it is not necessary for the sub-models to actually operate at the same scale.
For sub-models operating at a substantially different scale than the metamodel, scaling functions will be required; however, these functions can be simple while still contributing useful information to the metamodel (see Example 1). 
All predictions may therefore be integrated in a single Bayesian framework when evaluating the parameters of the metamodel (see Box 1, Fig. \ref{fig:diagram}).
We present and apply the framework here with two examples, and follow with a more formal presentation in Box 1.

%% ----------------------------------
%
%     EXAMPLE 1
%
%% ----------------------------------
\subsection*{Example 1: Adding experimental evidence for the fundamental niche to a species distribution model}
\ia{To keep the ms short, the 2 examples can be summarized a bit more and the details put in the Appendix}

@@

We consider data collected from a species' historical distribution, where the goal is to predict the distribution following a substantial reduction in precipitation. 
We will consider \(X_M\) (see Box 1, Fig. \ref{fig:diagram}) to be a vector of $n$ observations that takes the value of one to indicate presence and zero for absence:
\begin{equation}
X_M = \{x_{M,1}, x_{M,2}, \ldots, x_{M,n}\}
\end{equation}
We assume the data are relatively high-quality, providing coverage over a wide region and covering various climatic conditions (Fig. \ref{fig:ex1_sampling}a). 
The model is evaluated by relating these observations to environmental data \((D_M)\), which, for the sake of the example, consists of mean annual temperature $T_M$ and annual precipitation $P_M$. 
For simplicity, we formulate the metamodel \((\theta_M)\) as a simple logistic regression with a second order effect of temperature and precipitation. 
This naive model (i.e., the metamodel with no constraints from sub-models) predicts  occurrence probability (\(\psi_N\)) as a function of the environment:
%-----
\begin{equation}
\begin{aligned}
	\psi_N &= f\left(\theta_M, D_M \right) \\
	&= p \left (X_M = 1 \mid \theta_M, T_M, P_M \right) \\
	&=\text{logit}^{-1}\left( \mathbf{\Theta_M} \mathbf{D_M} \right)
\end{aligned}
\end{equation}
%-----
where \(\mathbf{\Theta_M}\) is the parameter vector of the model \(\theta_M\), \(\mathbf{D_M} \) is the covariate matrix (i.e., \(T_M, P_M\), with the first column taken to be unity to allow an intercept to be fit), and \(\text{logit}^{-1}\) is the inverse of the logit function.
We can fit this model in a Bayesian framework to allow for easy integration of models (as we will show later).
In this context, the goal of modeling is to estimate the probability distribution of \(\theta_M\), the model parameters, given the observed data \((X_M, T_M, P_M)\), which is given by the proportional form of Bayes's Theorem \citep[for readers unfamiliar with general concepts in Bayesian inference, a concise introduction can be found in ][]{Link2010}:
%-----
\begin{equation}
\label{eq:ex1_bayes}
	p\left (\theta_M \mid X_M,T_M,P_M \right ) \propto 
	p \left(X_M \mid \theta_M, T_M, P_M \right)
	p \left(\theta_M \right)
\end{equation}
%-----
where \(p\left(X_M \mid \theta_M, T_M, P_M \right)\) is often referred to as the \emph{likelihood} of the data (\(X_M\)) given the model (\(\theta_M\)), 
\wt{model parameters?}
\(p\left(\theta_M \right)\) is often referred to as the \emph{prior distribution} of \(\theta_M\), and the goal of modeling is to estimate \(p\left (\theta_M \mid X_M,T_M,P_M \right )\), the \emph{posterior distribution} of \(\theta_M\), which gives the probability that \(\theta_M\) takes particular values, given the observed data.

@@

We performed this procedure with our hypothetical datasets.
When comparing the three models (naive metamodel, mechanistic sub-model, and integrated metamodel), we observed extreme uncertainty in the first model when projecting beyond the range of the original data.
\ks{add ref to figure}
Unsurprisingly given the quality of the data, the sub-model was highly precise with respect to precipitation and not useful for temperature, thus providing a fairly strong constraint when producing the integrated model.
The result was an integrated prediction that reflected the shapes of both models and showed considerably reduced uncertainty (Fig. \ref{fig:ex1_precip}).
At the scale of the metamodel, considering both temperature and precipitation, we observed similar results, with reduced uncertainty in the predictions over the domain not covered by the presence-absence data (Fig. \ref{fig:ex1_map}).

@@

\subsection*{Example 2: Constraining an SDM using phenological information}
For the second example, we consider the problem of predicting the future distribution 
\wt{forecasting the potential distribution of a species}
of a species following climate change.
There is considerable interest in comparing correlative and mechanistic projections with respect to climate change \citep{Morin2009}, and correctly characterizing uncertainty is a critical aspect of this problem \citep{Cheaib2012}.
Despite being a relatively common application of \ac{SDM}s \citep{Guisan2005}, projecting models parameterized with modern climate data to future climate scenarios remains problematic \citep{Araujo2006}.
We used our metamodeling framework to inform a climate-based SDM with information obtained from Phenofit, a mechanistic model that predicts a species' probability of presence as a function of the suitability of the environment given the species' phenology \citep{Chuine2001, Morin2009}.
We use data from \citet{Morin2009} for sugar maple (\emph{Acer saccharum}), an economically and ecologically important species occurring in eastern North America.
Here we describe briefly the dataset, methods, and the results of the analysis (see Appendix S2 for data and scripts to reproduce the analysis).

>>>>

%%% available colors:
%%% red, green, blue, cyan, magenta, yellow, gray, drakfray, lightgray, brown, lime, olive, orange, pink, purple, teal, violet
%
\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
%\newcommand{\ib}[1]{\todo[color=magenta!40]{\textsuperscript{IB}#1}} 				
\renewcommand{\aa}[1]{\todo[color=violet!40]{\textsuperscript{Aitor A}#1}}
\newcommand{\ia}[1]{\todo[color=blue!20]{\textsuperscript{Isabelle A}#1}}
\newcommand{\db}[1]{\todo[color=violet!20]{\textsuperscript{Dominique B}#1}}
%\newcommand{\ab}[1]{\todo[color=teal!40]{\textsuperscript{Alyssa B}#1}}
%\newcommand{\fd}[1]{\todo[color=orange!30]{\textsuperscript{Frédérik D}#1}}
%\newcommand{\rd}[1]{\todo[color=orange!40]{\textsuperscript{Ronnie D}#1}}
\newcommand{\mjf}[1]{\todo[color=green!60]{\textsuperscript{Marie-Josée F.}#1}}
%\newcommand{\tf}[1]{\todo[color=yellow!60]{\textsuperscript{Tony F.}#1}}
\newcommand{\dm}[1]{\todo[color=lime!50]{\textsuperscript{Dan M.}#1}}
\newcommand{\ks}[1]{\todo[color=green!40]{\textsuperscript{Kevin S}#1}}
%\newcommand{\ns}[1]{\todo[backgroundcolor=red!100, linecolor=orange!100, bordercolor=cyan!100]{{\color{magenta!40}\textsuperscript{Strigul}#1}}}
\newcommand{\wt}[1]{\todo[color=orange!80]{\textsuperscript{Wilfried T}#1}}
\newcommand{\dg}[1]{\todo[color=pink!40]{\textsuperscript{DG}#1}}

@@

%TC:break abstract
\linenumbers
\begin{abstract}
% limited to 200 words
Current interest in modeling potential changes to species ranges have resulted in a great diversity of approaches to species distribution modeling.
However, in general, individual approaches include only a small subset of the available information about a species. 
Correlative models often predict presence or absence as a function of climate, but ignore  smaller-scale processes such as growth, fecundity, and dispersal.
Furthermore, different approaches often produce divergent predictions, with no simple method to reconcile them.
We present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods. 
Our method results in a metamodel that produces probabilistic estimates of species presence, reports uncertainty from both data and process error from all sub-models, and incorporates all of the information used as input for the original scale-specific sub-models.
We illustrate our approach using two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of the original data.
We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
Although implementation of our method can be technically challenging, we anticipate the results will be of wide interest, and we encourage collaboration between modelers and practitioners in applying our framework.
\end{abstract}

@@

Models of species range limits have wide applications and can play a large role in conservation biology, where they can be used as decision-making tools in biodiversity management \citep{Guisan2013}.
Due to large temporal and spatial scales as well as the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the ecological processes generating species range limits \citep{Wu1995, Levin1998}. 
Hence, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, evolution \citep{Jay2012}, hybrid zone dynamics \citep{Engler2013}, and forecasting the impacts of climate change on species distributions \citep{Thuiller2011, Blois2013, Thuiller2014}.

Despite the recognized potential of these models, difficulties in producing predictions with acceptable precision and low bias has limited their contributions \citep{Guisan2013}.
Two important constraints on modeling ecological processes can prevent models from making unbiased predictions with acceptable uncertainty: 
(1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
Poor data or incoherent models will result in predictions with unacceptably high uncertainty or precise but biased predictions.
In recent decades, however, modeling techniques have proliferated to take advantage of what is available in terms of both data and modeling platforms.
Model diversification also arises due to the diversity of processes generating species ranges \citep{Boulangeat2012}.
In the context of forecasting, mechanistic and correlative approaches represent quite different modes of inference, but both have strong limitations. 
Fine-scale mechanistic models may capture important ecological processes quite well, but due to their specificity, they may perform poorly when applied at the scale of species ranges.
For instance, biotic interactions (including trophic interactions) are usually not modeled mechanistically at the scale of species distributions because they are poorly known, or have not been recorded, despite being considered a key determinant of range limits \citep{Holt2005, Pigot2013}. 
On the other hand, coarse-scale correlative models statistically relating species occurrences to associated distributions of other variables have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
However, their predictions rely on the constancy of the relationships between occurrences and explanatory variables in time and space, implying that the selected variables are related to the processes limiting species ranges and that their correlations are constant for calibration and projection ranges in space and time \citep{Dormann2007}. 
Extrapolating beyond the scope of the original data (e.g., predicting ranges based on future climate) is therefore problematic, because nonlinearities in responses to novel combinations of the explanatory variables cannot be accommodated in models that do not simulate the underlying processes.
Furthermore, correlative models often do not explicitly incorporate ecological processes into their formulations, and when applied at coarse scales, they may miss important fine-scale processes, making a multi-scale approach more suitable when accurate forecasting is required \citep{Thuiller2013}.
On the other hand process-based models often require a variety of assumptions which may be poorly understood across space and time.
\mt{This last sentence could go, and the sentence before is actually a bit repetitive. Suggest reorganizing a bit to still end on this point, but reduce repetition}
\aa{you mentioned fine-scale mechanistic models and, on the other hand, coarse-scale correlative models. To talk now about process-based models sounds a bit bizarre and may make the reader wonder why are you mentioning them now.}
\ia{Finish abruptely here. Feels like one sentence is missing}
\ks{Feels like it is getting snuck in here…almost an after thought… I would recommend either removing it or developing it a tad further. }
%% ----------------------------------
%% ----------------------------------
%% ----------------------------------

\subsection*{Toward an integrated approach for modeling ranges}
To include knowledge of species from multiple sources and scales into distribution models, we present an application of hierarchical Bayesian methods that uses outputs from multiple model outputs to inform the results of the final model.
Our approach is an alternative to other approaches to multimodel inference that offers improved flexibility in incorporating data and theory from multiple scales and provides transparent estimation of uncertainty.
By integrating all available knowledge while producing a single prediction, our approach potentially mitigates the limitations inherent in each individual model, contributing to more robust predictions \citep{Pearson2003, Guisan2005, Araujo2006, Quillet2010}.
Some researchers have explicitly reproduced the hierarchical nature of ecological systems using hierarchical models \citep[e.g.][]{Royale2008, Boulangeat2012, Catterall2012, Strigul2012, Soranno2014}. 
Recent literature has also focused on the development of hybrid models, which allow for flexible combinations between mechanistic and phenomenological sub-models \citep{Gallien2010, Thuiller2013}. 
For instance, correlative models are used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while mechanistic approaches can include space and time dynamics and therefore account for dispersal processes \citep{Kearney2008}. 
This complementarity has been used to merge the two model types into hybrid models \citep[e.g.][]{Keith2008, Anderson2009, Smolik2010, Dormann2012, Boulangeat2014}. 
However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are poorly known and difficult to test \citep{Gallien2010}, and the sources and magnitudes of uncertainties are difficult to quantify. 
Moreover, hybrid approaches are not suitable for merging models that are based on the same processes but with different underlying hypotheses.

A commonly-employed alternative to hybrid models is the direct combination of predictions.
If models operate at the same scale and have compatible parameterizations, predictions can be combined using multi-model inference \citep[e.g., model averaging, ensemble forecasting;][]{Thuiller2004, Araujo2007}. 
If models are not comparable because they are based on different data or hypotheses, the use of convergent predictions from multiple models is a way to provide more robust projections \citep{Morin2009, Marmion2009, Cheaib2012, Serra-Diaz2013}.
However, the applicability of ensemble forecasts is limited; for example, it is not currently possible to evaluate the effects of convergent predictions on the total uncertainty of the outcomes, despite its potential importance in a prediction context.
A more comprehensive understanding of uncertainty can guide biodiversity management and prioritize future data collection by identifying the parameters that contribute most to variance in the model predictions \citep{McMahon2011}. 
Clearly, an approach that integrates predictions and uncertainty from multiple model types is needed \citep{Beck2012, Thuiller2013}. 
Such an approach would incorporate multiple modes of inference and integrate data from multiple sources and scales \citep{Levin1992, Peters2004, Thuiller2013}.

Here, we present a framework for modeling species range dynamics that aims to account for all the information available on a focal species, from a variety of sources and scales.
Contrary to hybrid and hierarchical models, the aim is not to link different sub-models into a single model, but to condition the predictions of a metamodel at the target scale (e.g., an entire species' range) with information from independent sub-models at a variety of spatial scales, allowing for as much flexibility as possible regarding the type of information included. 
We use the power and generality of a hierarchical Bayesian framework, which allows us to include multiple data sources and modes of inference \citep{VanOijen2005, Clark2006, Hobbs2011, Hartig2012}. 
Another advantage of Bayesian methods is that uncertainty in model outputs is intuitive to interpret and reflects uncertainty at all levels of organization \citep{Cressie2009, Hobbs2011}. 
We illustrate our approach with two examples.
First, we use simulated data to present a complete example of the application of the framework to multiple information sources that are relevant at very different scales.
In a second example, we apply the framework to combine predictions of a correlative model relating the range of sugar maple (\emph{Acer saccharum}), a widespread foundation tree species from eastern North America, to climate with the predictions of a phenological model, Phenofit \citep{Chuine2001}, with the goal of improving uncertainty estimation when projecting model predictions to future climate.

@@

The key idea of our approach is formulate a metamodel that integrates data at the same ecological scale as the desired predictions.
The parameters of this model can be constrained using the output of one or more sub-models that can operate at different scales than the metamodel.
As an example, the metamodel could take the form of a simple correlative \ac{SDM}, with one or more fine-scaled process-based sub-models.
Ideally, each sub-model incorporates different data types and hypotheses that specify underlying ecological mechanisms contributing to the model output.
The output of sub-models should provide additional predictions at a comparable ecological scale at which the metamodel operates, although it is not necessary for the sub-models to actually operate at the same scale.
For sub-models operating at a substantially different scale than the metamodel, scaling functions will be required; however, these functions can be simple while still contributing useful information to the metamodel (see Example 1). 
All predictions may therefore be integrated in a single Bayesian framework when evaluating the parameters of the metamodel (see Box 1, Fig. \ref{fig:diagram}).
We present and apply the framework here with two examples, and follow with a more formal presentation in Box 1.

%% ----------------------------------
%
%     EXAMPLE 1
%
%% ----------------------------------
\subsection*{Example 1: Adding experimental evidence for the fundamental niche to a species distribution model}

@@

We consider data collected from a species' historical distribution, where the goal is to predict the distribution following a substantial reduction in precipitation. 
We will consider \(X_M\) (see Box 1, Fig. \ref{fig:diagram}) to be a vector of $n$ observations that takes the value of one to indicate presence and zero for absence:
\begin{equation}
X_M = \{x_{M,1}, x_{M,2}, \ldots, x_{M,n}\}
\end{equation}
We assume the data are relatively high-quality, providing coverage over a wide region and covering various climatic conditions (Fig. \ref{fig:ex1_sampling}a). 
The model is evaluated by relating these observations to environmental data \((D_M)\), which, for the sake of the example, consists of mean annual temperature $T_M$ and annual precipitation $P_M$. 
For simplicity, we formulate the metamodel \((\theta_M)\) as a simple logistic regression with a second order effect of temperature and precipitation. 
This naive model (i.e., the metamodel with no constraints from sub-models) predicts  occurrence probability (\(\psi_N\)) as a function of the environment:
%-----
\begin{equation}
\begin{aligned}
	\psi_N &= f\left(\theta_M, D_M \right) \\
	&= p \left (X_M = 1 \mid \theta_M, T_M, P_M \right) \\
	&=\text{logit}^{-1}\left( \mathbf{\Theta_M} \mathbf{D_M} \right)
\end{aligned}
\end{equation}
%-----
where \(\mathbf{\Theta_M}\) is the parameter vector of the model \(\theta_M\), \(\mathbf{D_M} \) is the covariate matrix (i.e., \(T_M, P_M\), with the first column taken to be unity to allow an intercept to be fit), and \(\text{logit}^{-1}\) is the inverse of the logit function.
We can fit this model in a Bayesian framework to allow for easy integration of models (as we will show later).
In this context, the goal of modeling is to estimate the probability distribution of \(\theta_M\), the model parameters, given the observed data \((X_M, T_M, P_M)\), which is given by the proportional form of Bayes's Theorem \citep[for readers unfamiliar with general concepts in Bayesian inference, a concise introduction can be found in ][]{Link2010}:
%-----
\begin{equation}
\label{eq:ex1_bayes}
	p\left (\theta_M \mid X_M,T_M,P_M \right ) \propto 
	p \left(X_M \mid \theta_M, T_M, P_M \right)
	p \left(\theta_M \right)
\end{equation}
%-----
where \(p\left(X_M \mid \theta_M, T_M, P_M \right)\) is often referred to as the \emph{likelihood} of the data (\(X_M\)) given the model (\(\theta_M\)), \(p\left(\theta_M \right)\) is often referred to as the \emph{prior distribution} of \(\theta_M\), and the goal of modeling is to estimate \(p\left (\theta_M \mid X_M,T_M,P_M \right )\), the \emph{posterior distribution} of \(\theta_M\), which gives the probability that \(\theta_M\) takes particular values, given the observed data.

@@

We performed this procedure with our hypothetical datasets.
When comparing the three models (naive metamodel, mechanistic sub-model, and integrated metamodel), we observed extreme uncertainty in the first model when projecting beyond the range of the original data (Figs. \ref{fig:ex1_precip}a, \ref{fig:ex1_map}a, \ref{fig:ex1_map}b)
Unsurprisingly given the quality of the data, the sub-model was highly precise with respect to precipitation and not useful for temperature, thus providing a fairly strong constraint when producing the integrated model (Fig \ref{fig:ex1_precip}b).
The result was an integrated prediction that reflected the shapes of both models and showed considerably reduced uncertainty (Fig. \ref{fig:ex1_precip}).
At the scale of the metamodel, considering both temperature and precipitation, we observed similar results, with reduced uncertainty in the predictions over the domain not covered by the presence-absence data (Fig. \ref{fig:ex1_map}).

@@

\subsection*{Example 2: Constraining an SDM using phenological information}
For the second example, we consider the problem of forecasting the potential distribution of a species following climate change.
There is considerable interest in comparing correlative and mechanistic projections with respect to climate change \citep{Morin2009}, and correctly characterizing uncertainty is a critical aspect of this problem \citep{Cheaib2012}.
Despite being a relatively common application of \ac{SDM}s \citep{Guisan2005}, projecting models parameterized with modern climate data to future climate scenarios remains problematic \citep{Araujo2006}.
We used our metamodeling framework to inform a climate-based SDM with information obtained from Phenofit, a mechanistic model that predicts a species' probability of presence as a function of the suitability of the environment given the species' phenology \citep{Chuine2001, Morin2009}.
We use data from \citet{Morin2009} for sugar maple (\emph{Acer saccharum}), an economically and ecologically important species occurring in eastern North America.
Here we describe briefly the dataset, methods, and the results of the analysis (see Appendix S2 for data and scripts to reproduce the analysis).
